---
title: "Mystery Case Tracker Preprocessing"
subtitle: "This scheduled RMarkdown document queries the data model, performs preprocessing, updates the pinned datasources and restarts the Mystery Case Tracker app."
author:
  - name: Logan Wu^[Department of Health, logan.wu@health.vic.gov.au]
date: "Last refreshed `r format(lubridate::now(tzone='Australia/Melbourne'), '%d %B, %Y %H:%M')`"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
    fig_width: 8.64
toc-title: "On this page"
dh-navbar: yes
---

Note: This file is completely unused for the demo version of the app.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(dplyr)
library(tidyr)
library(stringr)
library(stringi)
library(sf)
library(odbc)
library(pins)
library(connectapi)
library(R.utils)
library(whichWithin) # devtools::install_github("HughParsonage/whichWithin")
library(data.table) # devtools::install_github("rstudio/connectapi")
library(xml2)
library(lobstr)

# Declare functions
'%ni%' <- Negate('%in%')

if (Sys.getenv("RSCONNECT_API_KEY") == "") {
  stop("No RStudio Connect API key supplied in .Renviron or with Sys.setenv(). Contact LW for assistance")
}
```

# Set up database connection

```{r make_connection}
# Add UID/PWD environment variables in RSConnect after publishing
message("Making connection")
conn <- dbConnect(
  odbc(),
  Driver = "ODBC Driver 17 for SQL Server",
  Server = "u-ause-ps-sql-s-hpb-rpt-01.database.windows.net",
  uid = Sys.getenv("PHESS_DW01_UID"),
  pwd = Sys.getenv("PHESS_DW01_PWD"),
  Database = "phess-dw01",
  encoding = "latin1",
  applicationintent = "readonly"
)
```

## Run SQL queries

Includes positive rows, rows with a `ClusterName`, and non-rejected rows within the past 60 days. This is a temporary solution to cope with the increased data size from permit holders.

```{r sql_classifications}
all_classifications = dbGetQuery(conn, "SELECT DISTINCT Classification FROM sf_derived.LineList")
all_classifications = all_classifications[!is.na(all_classifications)]
rejected_classifications = all_classifications %>%
  str_subset("eject")
```

```{r sql_linelist}
linelist <- dbGetQuery(conn, paste0("SELECT
                       CAST(ll.CaseNumber AS INT) AS CaseNumber,
                       ll.Classification,
                       pp.Latitude,
                       pp.Longitude,
                       ac.AddressLongitude__c AS AddressLongitude,
                       ac.AddressLatitude__c AS AddressLatitude
                       FROM V_LineList AS ll
                       LEFT JOIN V_Person as pp
                         ON ll.RecordID = pp.RecordID
                       LEFT JOIN Account as ac
                         ON pp.AccountID = ac.Id
                       WHERE ll.Classification NOT IN (", paste0("'", rejected_classifications, "'", collapse=", "), ")"))
linelist = linelist %>%
  mutate(
    # Set known default values to NA
    Longitude = ifelse(Longitude==-999 & Latitude==-999 |
                         Longitude==145 & Latitude%in%c(-39,-39.2), NA, Longitude),
    Latitude = ifelse(is.na(Longitude), NA, Latitude),
    AddressLongitude = ifelse(AddressLongitude==-999 & AddressLatitude==-999 |
                                AddressLongitude==145 & AddressLatitude%in%c(-39,-39.2), NA, AddressLongitude),
    AddressLatitude = ifelse(is.na(AddressLongitude), NA, AddressLatitude),
    
    # Coalesce coordinates and remove Account fields
    Longitude = coalesce(AddressLongitude, Longitude),
    Latitude = coalesce(AddressLatitude, Latitude),
    AddressLongitude = NULL,
    AddressLatitude = NULL
  )
```

```{r sql_clustersites}
clustersites <- dbGetQuery(conn, 
                           "SELECT
                           CAST(ExposureCaseNumber AS INT) AS ExposureCaseNumber,
                           DeclaredDate,
                           Latitude,
                           Longitude
                           FROM V_ClusterSites")
```

```{r sql_casemovement}
casemovement = dbGetQuery(conn,
                          "SELECT
                          CAST(ll.CaseNumber AS INT) AS CaseNumber,
                          cm.MovementDate,
                          cm.MovementTime,
                          cm.MovementPlace,
                          cm.FullCleanAddress AS 'Address',
                          COALESCE(lo.AddressLongitude__c, cm.Longitude) AS Longitude,
                          COALESCE(lo.AddressLatitude__c, cm.Latitude) AS Latitude,
                          cm.AddressType
                          FROM V_casemovement AS cm
                          INNER JOIN V_LineList AS ll
                            ON cm.RecordID = ll.RecordID
                          LEFT JOIN ContactEncounter AS ce
                            ON cm.Id = ce.Id
                          LEFT JOIN Location AS lo
                            ON ce.Location__c = lo.Id") %>%
  mutate_if(is.character, function(x) iconv(x, to="Latin1")) %>%
  # Set known default coordinates to NA
  mutate(Longitude = ifelse(Longitude==145 & Latitude%in%c(-39,-39.2), NA, Longitude),
         Latitude = ifelse(is.na(Longitude), NA, Latitude)) %>%
  # create unique casemovement location id from lat and lon:
  arrange(Latitude) %>%
  group_by(Longitude, Latitude) %>%
  mutate(CoordID = -cur_group_id()) %>%
  ungroup() %>%
  # remove the CoordID for invalid coordinates
  mutate(CoordID = ifelse(is.na(Longitude), NA, CoordID),
         Address = Address %>% str_to_title %>% str_replace("Vic(?![a-z])", "VIC"))

message("SQL queries complete")
```

- **Linelist**: `r scales::comma(nrow(linelist))` rows
- **Clustersites**: `r scales::comma(nrow(clustersites))` rows
- **Casemovements**: `r scales::comma(nrow(casemovement))` rows

## Proximity links between exposure sites and case movement locations

```{r location_proximity_links}
locations = bind_rows(
  clustersites %>% mutate(id=ExposureCaseNumber, Type="Exposure site"),
  casemovement %>% distinct(CoordID, .keep_all=T) %>% mutate(id=CoordID, Type="Risk history")
) %>%
  select(id, Longitude, Latitude) %>%
  drop_na(Longitude, Latitude)
setDT(locations)
setkey(locations, Latitude, Longitude)

# location date ranges
cooldown = 30 # a visit is active for 30 days
clustersites.daterange = clustersites %>%
  select(id = ExposureCaseNumber, start = DeclaredDate) %>%
  mutate(end = start + days(30))
casemovement.daterange = casemovement %>%
  select(id = CoordID, start = MovementDate) %>%
  mutate(end = start + days(30)) %>%
  drop_na(id) %>%
  group_by(id) %>%
  summarise(start = min(start), end = max(end))
locations.daterange = bind_rows(clustersites.daterange, casemovement.daterange)

# Add links with max distance
location_proximity_edges = which_within(locations$Latitude, 
                                        locations$Longitude,
                                        radius = "500 m",
                                        incl_dist = TRUE,
                                        id = locations$id) %>%
  mutate(distance = dist * 1000) %>%
  select(from = orig, to = dest, distance) %>%
  left_join(locations.daterange, by=c("from"="id")) %>%
  left_join(locations.daterange, by=c("to"="id")) %>%
  filter(start.x <= end.y & start.y <= end.x) %>% # req ranges to overlap
  select(from, to, distance)
```

Generated **`r nrow(location_proximity_edges)`** proximity edges between locations (exposure sites and risk history)

## Proximity links between people

- Max distance of 50m
- Requires at least one to be a confirmed case

```{r person_proximity_links}
persons = linelist %>%
  select(CaseNumber, Longitude, Latitude, Classification) %>%
  drop_na(Longitude, Latitude)
setDT(persons)
setkey(persons, Latitude, Longitude)

# Add links with very small max distance
person_proximity_edges = which_within(persons$Latitude, 
                                      persons$Longitude,
                                      radius = "50 m",
                                      incl_dist = TRUE,
                                      id = persons$CaseNumber) %>%
  select(from = orig, to = dest, distance=dist) %>%
  filter(from %in% persons$CaseNumber[persons$Classification=="Confirmed"] |
           to %in% persons$CaseNumber[persons$Classification=="Confirmed"])
```

Generated **`r nrow(person_proximity_edges)`** proximity edges between people where at least one is a confirmed case

# Write data to RSConnect board

```{r write_to_rsconnect}
# bind all proximity edges together then add to list with case movements
proximity_edges = bind_rows(location_proximity_edges, person_proximity_edges)
message("Board register complete")
data = list(casemovement = casemovement,
            proximity_edges = proximity_edges,
            all_classifications = all_classifications)

board_register_rsconnect(server = "https://rstudio.covid19.dhhs.vic.gov.au",
                         key = Sys.getenv("RSCONNECT_API_KEY"))
pin(data, "MysteryCaseTracker-Data", board="rsconnect")
```

Pinned **`r round(obj_size(data)/(1024^2), 1)`** megabytes
